Even-point Multi-loop Unitarity and its Applications: Exponentiation, Anomalies and Evanescence

### 2023-07-07 eventPointTwoLoop.tex
dunce -> ostrich

\begin{equation}\label{eq:NLSMDunce}
\begin{aligned}
\text{Cut}\left(\! \scaleIntBscalarsmall{1}{2}{3}{4} \! \right)&=\sum_{\text{states}}\mathcal{A}^{\text{NLSM}}_{(2|q_4\bar{q}_3|\bar{q}_1)}\mathcal{A}^{\text{NLSM}}_{(1|\bar{q}_4q_3|\bar{q}_2)}\mathcal{A}^{\text{NLSM}}_{(q_2|34|q_1)}
\\
&=C^{\text{Xbox}}_{([12]|34)}\left[\tau^{(3)}_1 \tau^{(2)}_3\tau^{(1)}_4+\tau^{(3)}_1 \tau^{(2)}_4\tau^{(1)}_3+\tau^{(3)}_2 \tau^{(2)}_3\tau^{(1)}_4+\tau^{(3)}_2 \tau^{(2)}_4\tau^{(1)}_3\right]
\\
&-\left(C^{\text{2box}}_{(12|34)}\tau^{(3)}_1+C^{\text{2box}}_{(12|43)}\tau^{(3)}_2\right) \left[ \tau^{(2)}_3\tau^{(1)}_3+ \tau^{(2)}_4\tau^{(1)}_4\right] \,.
\end{aligned}
\end{equation}
The kinematic variables $\tau^{(i)}_j $ and $\tau_{ij}$ are the same, except we have made the replacement $p_i \rightarrow q_i$. The full two-loop NLSM amplitude can thus be computed by summing over the distinct labels of the resulting integrals, each weighted by suitable internal symmetry factors:
\begin{equation}
\label{eq:NLSM2loop}
\mathcal{A}^{\text{NLSM}}_{\text{2-loop}}=\frac{1}{4}\left[ \scaleIntCsmall \right]+ \frac{1}{2}\left[\scaleIntBscalarsmall{1}{2}{3}{4}+\scaleIntBscalarsmall{4}{3}{2}{1}\right]+\text{cyc}(2,3,4)
\end{equation}
We stress that the integrals appearing above are complicated by tensor integrals with many powers of loop momenta. However, as noted in the introduction, since both integrals are recursively one-loop, we can again apply a two-loop generalization of Passarino-Veltman for both the bubble and triangle integrals, stated in \eqn{eq:bubRed} and \eqn{eq:triRed}, respectively. With this, we proceed to the next step in the calculation.

@@

\begin{equation}\label{eq:NLSMdunce}

@@

We have added the additional factor of $(l_2+k_1)^2$ in the definition of $[I_3\circ I_2]^D(k_{12})$ to simplify the final expression in \eqn{eq:NLSMdunce}. We will use this as our basis integral throughout the text. This two-loop integral can be evaluated analytically using the $D$-dimensional single scale bubble and triangle integral expressions in \eqn{eq:integrals}. Thus, we have completely reduced the full two-loop NLSM amplitude to a linear combination of scalar integrals, $I_2^D(k_{ij})$ and $[I_3\circ I_2]^D(k_{ij})$. Now we will proceed to the final step of evaluating the amplitude in particular dimensions.
\paragraph{\textbf{Integration}} While there is plenty of physics to be extracted from the combination of \eqn{eq:NLSM2bub} and \eqn{eq:NLSMdunce}, we will focus on the behavior of the leading divergence for $D=2-2\epsilon$ in the planar limit where $N_c \rightarrow \infty$. In this limit, the cross-box is subleading in the $N_c$ expansion, and can be neglected. In a trace basis, the double-box can be expanded as follows:

@@

The analogous divergences for $D=4-2\epsilon$ can be extracted directly from the integral expressions in \eqn{eq:integrals}. Plugging these values into \eqn{eq:NLSM2bub} and \eqn{eq:NLSMdunce}, we can extract the full leading logarithmic divergence from the full color dressed amplitude in \eqn{eq:NLSM2loop}. Since we are interested in the large $N_c$ limit, we will define a color-ordered quantity $A^{2-2\epsilon}_{(1234)}$, 

@@

As in the previous section, exposed legs implicitly sum over internal states. The grey blobs above represent $D$-dimensional $t_8F^4$ operator insertions from the four-point BI tree amplitudes, which we labelled as $\mathcal{M}^{\gamma\gamma\gamma\gamma}_{(1234)}$. We give the internal loop momenta the same internal labels as we did the pions in \eqn{eq:NLSM2Bub} and \eqn{eq:NLSMDunce}, giving us the following integrals to evaluate for two-loop pure BI, 

@@

\subsubsection{Two-loop $\mathcal{N}=4$ DBIVA} \label{sec:DBIvDC2loop}
At two-loop it sufficient to show that replacing the color factors reproduces the integrals we found above with the unitarity construction. Starting with the NLSM integrands in \eqn{eq:NLSM2bub} and \eqn{eq:NLSMdunce}, we make the same replacement of color factors post integration as was done at one-loop.  We carry this out first for the double-bubble, and find that the full $D$-dimensional NLSM integral simplifies dramatically:
\begin{equation}
\begin{aligned}
\scaleIntCsmall \bigg|^{C_g \rightarrow N^{\mathcal{N}=4}_g}  \!\!&= N^{(12|34)}_{\mathcal{N}=4}\left[\frac{(s_{12}I_2^D(k_{12}))^2}{2}\left(\frac{s_{14}-s_{13}}{(D-1)^2}+s_{12}\right)\right]
\\
&+N^{(12|43)}_{\mathcal{N}=4}\left[\frac{(s_{12}I_2^D(k_{12}))^2}{2}\left(\frac{s_{13}-s_{14}}{(D-1)^2}+s_{12}\right)\right]
\\
&=s_{12}^4 (I_2^{D}(k_{12}))^2(t_8 F^4)^{\text{(max)}}_{(1234)}   \equiv \scaleIntCtune{}{hgreen}{}{hgreen} 
\end{aligned}
\end{equation}
The cancellation between dimension dependent factors is even more startling for the ostrich-diagram integral:
\begin{equation}
\begin{aligned}
\scaleIntBscalarsmall{1}{2}{3}{4}\bigg|^{C_g \rightarrow N^{\mathcal{N}=4}_g} \!\!&= N^{(12|34)}_{\mathcal{N}=4}\frac{s_{12}}{6} \left[\frac{(D-1)(D-4)s_{14}+2(D-2)^2 s_{13}}{(D-1)(4-3D)}\right][I_3\circ I_2]^D(k_{12})
\\
&+N^{(12|34)}_{\mathcal{N}=4}\frac{s_{12}}{6} \left[\frac{(D-1)(D-4)s_{13}+2(D-2)^2 s_{14}}{(D-1)(4-3D)}\right][I_3\circ I_2]^D(k_{12})
\\
&+N^{([12]|34)}_{\mathcal{N}=4}\left[\frac{s_{12}^2}{6}\frac{D+1}{D-1}\right][I_3\circ I_2]^D(k_{12})
\\
&=\frac{s_{12}^3}{3}[I_3\circ I_2]^{D}(k_{12})(t_8 F^4)^{\text{(max)}}_{(1234)} \equiv \scaleIntBtune{}{}{}{}{hgreen}{hgreen}{hgreen}{hgreen} 
\end{aligned}
\end{equation}
This alone is sufficient to demonstrate the validity of the double-copy, as these integrated quantities are exactly the same as those produced via generalized unitarity of $\mathcal{N}=4$ DBIVA at two-loop in \eqn{eq:N42bub} and \eqn{eq:N4ostrich}. We emphasize that based on our analysis, this consistency of the double-copy at two loop appears to hold in {any} dimension, as all the $D$-dependent prefactors drop out upon replacing the gauge theory color factors with $\mathcal{N}=4$ numerators. Considering the consistency of these two construction, this also serves as strong evidence for the existence of color-dual representation for two-loop pion integrands. We leave identifying such valid representation as an enticing future direction worth investigation.

@@

\paragraph{\textbf{Exponentiation}} In \eqn{eq:NLSM2bub} and \eqn{eq:NLSMdunce} we computed in general dimension, $D$, the two contributions to NLSM two-loop amplitudes. Evaluating these diagrams in $D=2-2\epsilon$, and taking the planar limit, $N_c \rightarrow \infty$, we found in \eqn{eq:expNLSM} that the IR divergences exponentiate through two-loop. This non-trivial property is found in theories that are conjectured to be integrable \cite{Shankar:1977cm,Zamolodchikov:1977nu,Zamolodchikov:1978xm, Komatsu:2019hgc}, like $\mathcal{N}=4$ super-Yang-Mills \cite{Anastasiou:2003kj,Bern:2005iz}. In future work, we hope to study whether this iterative structure can be further applied to the scale dependent logarithms present at multi-loop order in the theory. 

>>>>

\begin{equation}\label{eq:NLSMostrich}
\begin{aligned}
\text{Cut}\left(\! \scaleIntBscalarsmall{1}{2}{3}{4} \! \right)&=\sum_{\text{states}}\mathcal{A}^{\text{NLSM}}_{(2|q_4\bar{q}_3|\bar{q}_1)}\mathcal{A}^{\text{NLSM}}_{(1|\bar{q}_4q_3|\bar{q}_2)}\mathcal{A}^{\text{NLSM}}_{(q_2|34|q_1)}
\\
&=C^{\text{Xbox}}_{([12]|34)}\left[\tau^{(3)}_1 \tau^{(2)}_3\tau^{(1)}_4+\tau^{(3)}_1 \tau^{(2)}_4\tau^{(1)}_3+\tau^{(3)}_2 \tau^{(2)}_3\tau^{(1)}_4+\tau^{(3)}_2 \tau^{(2)}_4\tau^{(1)}_3\right]
\\
&-\left(C^{\text{2box}}_{(12|34)}\tau^{(3)}_1+C^{\text{2box}}_{(12|43)}\tau^{(3)}_2\right) \left[ \tau^{(2)}_3\tau^{(1)}_3+ \tau^{(2)}_4\tau^{(1)}_4\right] \,.
\end{aligned}
\end{equation}
The kinematic variables $\tau^{(i)}_j $ and $\tau_{ij}$ are the same, except we have made the replacement $p_i \rightarrow q_i$. The full two-loop NLSM amplitude can thus be computed by summing over the distinct labels of the resulting integrals, each weighted by suitable internal symmetry factors:
\begin{equation}
\label{eq:NLSM2loop}
\mathcal{A}^{\text{NLSM}}_{\text{2-loop}}=\frac{1}{4}\left[ \scaleIntCsmall \right]+ \frac{1}{2}\left[\scaleIntBscalarsmall{1}{2}{3}{4}+\scaleIntBscalarsmall{4}{3}{2}{1}\right]+\text{cyc}(2,3,4)
\end{equation}
We stress that the integrals appearing above are complicated by tensor integrals with many powers of loop momenta. However, as noted in the introduction, since both integrals are recursively one-loop, we can again apply a two-loop generalization of Passarino-Veltman for both the bubble and triangle integrals, stated in \eqn{eq:bubRed} and \eqn{eq:triRed}, respectively. With this, we proceed to the next step in the calculation.

@@

\begin{equation}\label{eq:NLSMostrich}

@@

We have added the additional factor of $(l_2+k_1)^2$ in the definition of $[I_3\circ I_2]^D(k_{12})$ to simplify the final expression in \eqn{eq:NLSMostrich}. We will use this as our basis integral throughout the text. This two-loop integral can be evaluated analytically using the $D$-dimensional single scale bubble and triangle integral expressions in \eqn{eq:integrals}. Thus, we have completely reduced the full two-loop NLSM amplitude to a linear combination of scalar integrals, $I_2^D(k_{ij})$ and $[I_3\circ I_2]^D(k_{ij})$. Now we will proceed to the final step of evaluating the amplitude in particular dimensions.
\paragraph{\textbf{Integration}} While there is plenty of physics to be extracted from the combination of \eqn{eq:NLSM2bub} and \eqn{eq:NLSMostrich}, we will focus on the behavior of the leading divergence for $D=2-2\epsilon$ in the planar limit where $N_c \rightarrow \infty$. In this limit, the cross-box is subleading in the $N_c$ expansion, and can be neglected. In a trace basis, the double-box can be expanded as follows:

@@

The analogous divergences for $D=4-2\epsilon$ can be extracted directly from the integral expressions in \eqn{eq:integrals}. Plugging these values into \eqn{eq:NLSM2bub} and \eqn{eq:NLSMostrich}, we can extract the full leading logarithmic divergence from the full color dressed amplitude in \eqn{eq:NLSM2loop}. Since we are interested in the large $N_c$ limit, we will define a color-ordered quantity $A^{2-2\epsilon}_{(1234)}$, 

@@

As in the previous section, exposed legs implicitly sum over internal states. The grey blobs above represent $D$-dimensional $t_8F^4$ operator insertions from the four-point BI tree amplitudes, which we labelled as $\mathcal{M}^{\gamma\gamma\gamma\gamma}_{(1234)}$. We give the internal loop momenta the same internal labels as we did the pions in \eqn{eq:NLSM2Bub} and \eqn{eq:NLSMostrich}, giving us the following integrals to evaluate for two-loop pure BI, 

@@

\subsubsection{Two-loop $\mathcal{N}=4$ DBIVA} \label{sec:DBIvDC2loop}
At two-loop it sufficient to show that replacing the color factors reproduces the integrals we found above with the unitarity construction. Starting with the NLSM integrands in \eqn{eq:NLSM2bub} and \eqn{eq:NLSMostrich}, we make the same replacement of color factors post integration as was done at one-loop.  We carry this out first for the double-bubble, and find that the full $D$-dimensional NLSM integral simplifies dramatically:
\begin{equation}
\begin{aligned}
\scaleIntCsmall \bigg|^{C_g \rightarrow N^{\mathcal{N}=4}_g}  \!\!&= N^{(12|34)}_{\mathcal{N}=4}\left[\frac{(s_{12}I_2^D(k_{12}))^2}{2}\left(\frac{s_{14}-s_{13}}{(D-1)^2}+s_{12}\right)\right]
\\
&+N^{(12|43)}_{\mathcal{N}=4}\left[\frac{(s_{12}I_2^D(k_{12}))^2}{2}\left(\frac{s_{13}-s_{14}}{(D-1)^2}+s_{12}\right)\right]
\\
&=s_{12}^4 (I_2^{D}(k_{12}))^2(t_8 F^4)^{\text{(max)}}_{(1234)}   \equiv \scaleIntCtune{}{hgreen}{}{hgreen} 
\end{aligned}
\end{equation}
The cancellation between dimension dependent factors is even more startling for the ostrich-diagram integral:
\begin{equation}
\begin{aligned}
\scaleIntBscalarsmall{1}{2}{3}{4}\bigg|^{C_g \rightarrow N^{\mathcal{N}=4}_g} \!\!&= N^{(12|34)}_{\mathcal{N}=4}\frac{s_{12}}{6} \left[\frac{(D-1)(D-4)s_{14}+2(D-2)^2 s_{13}}{(D-1)(4-3D)}\right][I_3\circ I_2]^D(k_{12})
\\
&+N^{(12|34)}_{\mathcal{N}=4}\frac{s_{12}}{6} \left[\frac{(D-1)(D-4)s_{13}+2(D-2)^2 s_{14}}{(D-1)(4-3D)}\right][I_3\circ I_2]^D(k_{12})
\\
&+N^{([12]|34)}_{\mathcal{N}=4}\left[\frac{s_{12}^2}{6}\frac{D+1}{D-1}\right][I_3\circ I_2]^D(k_{12})
\\
&=\frac{s_{12}^3}{3}[I_3\circ I_2]^{D}(k_{12})(t_8 F^4)^{\text{(max)}}_{(1234)} \equiv \scaleIntBtune{}{}{}{}{hgreen}{hgreen}{hgreen}{hgreen} 
\end{aligned}
\end{equation}
This alone is sufficient to demonstrate the validity of the double-copy, as these integrated quantities are exactly the same as those produced via generalized unitarity of $\mathcal{N}=4$ DBIVA at two-loop in \eqn{eq:N42bub} and \eqn{eq:N4ostrich}. We emphasize that based on our analysis, this consistency of the double-copy at two loop appears to hold in {any} dimension, as all the $D$-dependent prefactors drop out upon replacing the gauge theory color factors with $\mathcal{N}=4$ numerators. Considering the consistency of these two construction, this also serves as strong evidence for the existence of color-dual representation for two-loop pion integrands. We leave identifying such valid representation as an enticing future direction worth investigation.

@@

\paragraph{\textbf{Exponentiation}} In \eqn{eq:NLSM2bub} and \eqn{eq:NLSMostrich} we computed in general dimension, $D$, the two contributions to NLSM two-loop amplitudes. Evaluating these diagrams in $D=2-2\epsilon$, and taking the planar limit, $N_c \rightarrow \infty$, we found in \eqn{eq:expNLSM} that the IR divergences exponentiate through two-loop. This non-trivial property is found in theories that are conjectured to be integrable \cite{Shankar:1977cm,Zamolodchikov:1977nu,Zamolodchikov:1978xm, Komatsu:2019hgc}, like $\mathcal{N}=4$ super-Yang-Mills \cite{Anastasiou:2003kj,Bern:2005iz}. In future work, we hope to study whether this iterative structure can be further applied to the scale dependent logarithms present at multi-loop order in the theory. 
